Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}

### 2023-06-09 paper/paper.tex
#22 add levine explanation

The European state system was forged in conflict. No other continent at any time experienced more war than Europe in the early modern era (\citealp{voigtlnder2013}, p. 174). Wars brought death and devastation to the societies affected by them, but their effects also manifested on the political map --- they play out in a diverse ecosystem of polities that come and go and change size over time. Borders get redrawn and territory changes hands. One can indeed imagine this as an evolutionary process: states compete for resources and powerful states eventually absorb weaker ones. As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while unfavourable traits are doomed to disappear \citep{levine2013, levine2021}.

@@

@@

My story is not about population but institutions (although it is compatible with Malthusianism). Explaining long-run growth with institutional factors has a long tradition in economics: North, Acemoglu, property rights, civil law, middle class. The key institutional change brought about by conflict was fiscal capacity \citep{tilly1985, dincecco2012}, or the state's ability to raise taxes. A newer strand of research focuses on fiscal capacity:  \cite{cantoni2023}

Theory of wars and states: \cite{tilly1985} argues that war was the main reason for governments to invest in fiscal capacity. The formal theory of conflict-driven evolution is from \cite{levine2013}.

\cite{diamond1997} and \cite{landes1969, landes2006} have argued for the economic benefits of political fragmentation and state competition. \cite{diamond1997} asserts that geography 
remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history. \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system.

@@

\hspace{0.5cm}

>>>>

The European state system was forged in conflict. No other continent at any time experienced more war than Europe in the early modern era (\citealp{voigtlnder2013}, p. 174). Wars brought death and devastation to the societies affected by them, but their effects also manifested on the political map --- they play out in a diverse ecosystem of polities that come and go and change size over time. Borders get redrawn and territory changes hands. One can indeed imagine this as an evolutionary process: states compete for resources and powerful states eventually absorb weaker ones. As a result, traits that increase state power --- the ability to win conflicts against rivals --- get selected for, while unfavourable traits are doomed to disappear \citep{levine2021}.

@@

I build my paper on the theory of conflict driven evolution, outlined above, developed by \cite{levine2013, levine2021, levine2022}. Their model explains why hegemonies arose in East Asia while balances of power could prevail in Europe: unlike China, the major powers of continental Europe were frequently preyed upon by strong outsiders --- the Vikings, the Swedes, and the English --- which prevented any one state from growing too strong (\citealp{levine2021}, p. 439).
However, it focuses more on these aggregate patterns than on the traits that cause states to succeed.
\footnote{The basic idea of Darwinian selection has previously been borrowed by economists, though - for example \cite{galor2002} do the literal version, where humans are subject to selection and accumulate favourable traits over the course of history. More infamously, \cite{clark2007} employs it in an attempt to explain why the Industrial Revolution took place in Europe, arguing that the rich out-bred the poor and passed favourable traits on to their offspring. in Europe but not elsewhere.}

@@

My story is not about population but institutions (although it is compatible with Malthusianism). A longstanding tradition in economics emphasises the role of institutional factors in long-run growth (e.g. \cite{north1970}, \cite{delong1993}, \cite{ajr2001}). From this vast literature one theme is especially relevant here: the institution of fiscal capacity, or the state's ability to collect taxes. \cite{tilly1985} argues that the main reason for states to collect taxes is to be able to wage war. In other words, fiscal capacity and state power are tightly linked --- if conflict selects for state power, then it also selects for fiscal capacity. \cite{cantoni2023} show that drove state consolidation and played a role in growth. \cite{dincecco2012}.

\cite{diamond1997} and \cite{landes1969, landes2006} have argued for the economic benefits of political fragmentation and state competition. \cite{diamond1997} asserts that geography 
remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history. \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled the spread of innovation in medieval China would not have survived long in a more fragmented state system. Fragmentation also plays a central role in \cite{cervellati2022}'s theory.

@@
